---
title: "DA5020.Practicum 1.Chunsheng.Fang"
output: 
  pdf_document: default
  html_document: default
date: "2024-10-010"
---
install.packages("ggplot2")


# Part 1
## Problem 1: 1. Create dataframe
```{r}
doctor_type <- c("PCP","Psychiatrist", "Surgeon", "Anesthesia")
doctor_lastname <- c("Smith","Dame", "Jones", "Zayas")
location<- c("MA", "ME","NH","VT")
AVG_Rating <- c(7, 9, 8, 9)
Doctor <-data.frame(doctor_type, doctor_lastname, location, AVG_Rating)
Doctor

```
## Problem 2: 2. Row and Column Selection
```{r}
Doctor[1,2]
Doctor[2:4,]
Doctor[,4]

```
## Problem 3:Use 2 variables in mtcars to create scatterplot
```{r}
library(ggplot2)
names(mtcars)
ggplot(mtcars, aes(x = mpg, y = wt)) +
geom_point(aes(colour = factor(gear)),size = 3) +
xlab("Miles per gallon") +
ylab("Weight") +
ggtitle("Relationship between MPG and Weight")

```
### I picked these variables because they represent two important factors that can impact a car’s performance: fuel efficiency and weight. The relationship between these variables can help us understand the trade-offs that car manufacturers make when designing vehicles, as well as provide insights into how these factors affect a car’s overall performance.

## Problem 4: Summary of mtcars and calculate pearson coefficient calculation and explain the purpose
```{r}
summary(mtcars)
cor(log10(mtcars$mpg),mtcars$hp,use = "complete.obs")

```
I chose these variables because they are both important factors in a car’s performance. mpg is fuel efficiency, while hp is the power of the car’s engine. By examining the correlation between these variables, we can understand how fuel efficiency and engine power are related.
The purpose of the Pearson coefficient of correlation is to measure the linear relationship between two variables. The Pearson coefficient is a standardized measure that ranges from -1 to 1, with 0 indicating no correlation and values closer to -1 or 1 indicating a stronger correlation. This coefficient is useful for examining the strength and direction of the relationship between two variables
In this case it’s a negative correlation and suggests that cars with more powerful engines tend to have lower fuel efficiency

#PART 2
## Problem 1
```{r}
# Install necessary packages
if (!require(httr)) install.packages("httr")
if (!require(readr)) install.packages("readr")

# Load libraries
library(httr)
library(readr)

# URL of the CSV file
PATH <- "/Users/chunsheng/Desktop/chemical-dependence-treatment-program-admissions-beginning-2007-1.csv"

ny.cd_df <-read.csv(PATH)

head(ny.cd_df)
```

## Problem 2
```{r}
# Set CRAN mirror
options(repos = c(CRAN = "https://cran.rstudio.com/"))
if (!require(dplyr))install.packages("dplyr")
library(dplyr)
if (!require(psych))install.packages("psych")
library(psych)
dim(ny.cd_df)
glimpse(ny.cd_df)
summary(ny.cd_df)
describe(ny.cd_df[1:7])
sum(is.na(ny.cd_df))
ny.cd_df <- ny.cd_df %>%
rename("County_of_Program_Location" = "County.of.Program.Location",
"Service_Type" = "Service.Type",
"Program_Category" = "Program.Category",
"Age_Group" = "Age.Group",
"Primary_Substance_Group" = "Primary.Substance.Group")
Adm <- ny.cd_df$Admissions
#summary of admission column
summary(Adm)
#mean of admission column
meann_Adm <- mean(Adm)
#Standard deviation of admission column
Std.Dev <- sd(Adm)
z <-abs((Adm - meann_Adm)/Std.Dev)
outliers_Adm <-which(z>3)
length(outliers_Adm)
```
## Problem 3:Structure the data and create four tibbles
```{r}
county <- tibble(county_code = c("AL","CA","CN","DE","FR","LE","MG","ON","OL","NY","SL","SY","SV","WR","AG","CY","CL","DU","FU","HE","LI","NA","OD","OS","RE","SA","SE","TI","WS","NY","CH","CO","ER", "GE","JE","MA","NY","OT","OG","NY","SC","ST","TO","WA","BM","CM","CR","ES","GR","NY","MO","NI","OR","PU","RO","SH","SU","UL","WE","WY","YA"),
County_of_Program_Location = c("Albany","Cattaraugus","Chenango","Delaware","Franklin","Lewis","Montgomery","Oneida","Orleans","Queens","Saint Lawrence",
"Schuyler","Sullivan","Warren","Allegany","Cayuga","Clinton","Dutchess","Fulton","Herkimer","Livingston","Nassau",
"Onondaga","Oswego","Rensselaer","Saratoga","Seneca","Tioga","Washington","Bronx","Chautauqua","Columbia","Erie","Genesee","Jefferson","Madison",
"New York","Ontario","Otsego","Richmond","Schenectady","Steuben","Tompkins","Wayne","Broome","Chemung","Cortland","Essex","Greene","Kings","Monroe",
"Niagara","Orange","Putnam","Rockland","Schoharie","Suffolk","Ulster","Westchester","Wyoming","Yates"))
county %>% distinct(County_of_Program_Location,.keep_all = TRUE)


Programcategory <- tibble(program_code = c("CR","IN", "OTP","Res", "OP", "SP" ),
Program_Category = c("Crisis", "Inpatient", "Opioid Treatment Program", "Residential", "Outpatient", "Specialized"))
Programcategory %>% distinct(program_code,.keep_all = TRUE)

ny.cd_df %>% distinct(Primary_Substance_Group)

Prim_sub_group <- tibble(substance_code = c("H","Al", "M","C", "Oth", "OP", "No"),
Primary_Substance_Group = c("Heroin", "Alcohol", "Marijuana", "Cocaine", "All Others", "Other Opioids", "None"))
Prim_sub_group %>% distinct(substance_code,.keep_all = TRUE)

ny.cd_df_new <- merge(x = ny.cd_df, y = county, by = "County_of_Program_Location",all= T)
ny.cd_df_new <- merge(x = ny.cd_df_new, y = Programcategory, by = "Program_Category",all= T)
ny.cd_df_new <- merge(x = ny.cd_df_new, y = Prim_sub_group, by = "Primary_Substance_Group",all= T)

# Select the required columns from new dataframe
Admission_tibble <- ny.cd_df_new %>% select("Year", "county_code", "program_code", "Service_Type", "Age_Group", "substance_code", "Admissions")
# Convert to a tibble
Admission <- as_tibble(Admission_tibble)
# View the resulting tibble
Admission



```
## Problem 4: Admission Data
```{r}
annualAdmissions <- function(data) {
# Aggregate admissions by year
admissions_by_year <- Admission %>%
group_by(Year)%>%
summarise("Admissions" = sum(Admissions))
# Add annotation for year with highest number of admissions
max_year <- admissions_by_year[which.max(admissions_by_year$Admissions), "Year"]
max_admissions <- max(admissions_by_year$Admissions)
# Create line chart
ggplot(admissions_by_year, aes(x = Year, y = Admissions)) +
geom_line() +
ylab("Number of Admissions") +
ggtitle("Total Admissions in New York State") }
# Example usage
annualAdmissions(Admission)


```
## Problem 5
```{r}

# Calculate percentage of admissions by county
ny_data <- ny.cd_df %>%
group_by(County_of_Program_Location) %>%
summarize(Percent_Admissions = sum(Admissions)/nrow(ny.cd_df) * 100)
# Sort data by percentage of admissions in descending order
ny_data <- ny_data %>%
arrange(desc(Percent_Admissions))
ny_data

# Select top 5 counties
top_counties <- head(ny_data, 5)
# Create bar chart of top 5 counties by percentage of admissions
ggplot(top_counties, aes(x = County_of_Program_Location, y = Percent_Admissions)) +
geom_bar(stat = "identity", fill = "steelblue") +
labs(x = "County", y = "Percent Admissions", title = "Top 5 Counties by Percentage of Admissions")
```
## Problem 6
```{r}
# Corrected syntax for checking if stringr is installed and installing if not
if (!require(stringr)) {
  install.packages("stringr")
}

library(stringr)
rehab_facility <- ny.cd_df %>%
filter(str_detect(Service_Type, "(R|r)ehab"))
substance_frequency <- rehab_facility %>%
group_by(Age_Group) %>%
count(Primary_Substance_Group)
substance_frequency

ggplot(substance_frequency, aes(x = Primary_Substance_Group, y = n, fill = Primary_Substance_Group)) +
geom_bar(stat = "identity") +
labs(x = "Age Group", y = "Frequency") +
facet_wrap("Age_Group", scales = "free_y", strip.position = "bottom") +
theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```