---
title: "DA5020.A5.Chunsheng.Fang"
output: 
  html_document: default
  pdf_document: default
date: "2024-10-06"
---



# Problem 1
# Load XML file directly from URL
```{r}
# Set a CRAN mirror explicitly (this avoids the non-interactive error)
options(repos = c(CRAN = "https://cran.rstudio.com/"))

# Install necessary packages if not already installed
if (!require(RCurl)) install.packages("RCurl")
if (!require(xml2)) install.packages("xml2")
if (!require(dplyr)) install.packages("dplyr")

# Load the libraries
library(RCurl)
library(xml2)
library(dplyr)

# URL of the XML file
url <- "https://www.senate.gov/general/contact_information/senators_cfm.xml"

# Download the XML content using RCurl
xml_content <- getURL(url)

# Parse the XML content
xml_parsed <- read_xml(xml_content)

# Extract relevant information from the XML into a data frame
# We need to extract each senator's details from the XML nodes
senators <- xml_find_all(xml_parsed, "//member")

# Create a data frame by extracting specific details like first_name, last_name, state, etc.
senator_data <- tibble(
  first_name = xml_text(xml_find_all(senators, ".//first_name")),
  last_name = xml_text(xml_find_all(senators, ".//last_name")),
  state = xml_text(xml_find_all(senators, ".//state")),
  party = xml_text(xml_find_all(senators, ".//party")),
  phone = xml_text(xml_find_all(senators, ".//phone")),
  address = xml_text(xml_find_all(senators, ".//address"))
)

# Display the dimensions of the data frame (rows and columns)
print(dim(senator_data))

# View the first few rows of the data
print(head(senator_data))


```
# Problem 2
#Construct a regular expression (regex) to extract only the first, last names, and party (D,R,I) of each senator
```{r}
# Regular expression to clean names
senator_data <- senator_data %>%
  mutate(
    # Clean first_name to remove middle initials
    first_name = sub("^(\\w+)(?:\\s+[A-Z]\\.)?\\s*.*$", "\\1", first_name),
    # Clean last_name to remove suffixes
    last_name = sub("^(\\w+)(?:\\s+(Jr|Sr|III|IV|II)\\.?)?$", "\\1", last_name),
    # Standardize party to D, R, I
    party = toupper(substr(party, 1, 1))
  ) %>%
  select(first_name, last_name, party, state, phone, address)

# Display the dimensions of the cleaned data frame (rows and columns)
print(dim(senator_data))

# View the first few rows of the cleaned data
print(head(senator_data))
```
# Problem 3
```{r}

# Clean party names and make them more readable
senator_data <- senator_data %>%
  mutate(
    party = case_when(
      party == "D" ~ "Democratic Party",
      party == "R" ~ "Republican Party",
      party == "I" ~ "Independent Party",
      TRUE ~ party
    )
  )

# Define the function to display senators by state
senatorsByState <- function(state_abbr) {
  # Filter the senator_data for the specified state
  state_senators <- senator_data %>%
    filter(state == state_abbr)

  # Check if there are any senators for the given state
  if (nrow(state_senators) > 0) {
    # Create a message with the senator names and parties
    senator_list <- apply(state_senators, 1, function(row) {
      paste(row["first_name"], row["last_name"], "(", row["party"], ")", sep = "")
    })
    
    # Join the names into a single string
    senators_message <- paste(senator_list, collapse = " and ")

    # Print the message
    message <- paste("The senators for", state_abbr, "are:", senators_message)
    print(message)
  } else {
    print(paste("No senators found for state:", state_abbr))
  }
}

# Example usage
senatorsByState("MA")  # You can change "MA" to any state abbreviation you want to check.


```
# Problem 4
```{r}
# Install the necessary package if not already installed
if (!require(readr)) {
  install.packages("readr")
}
library(readr)
# Install necessary packages if not already installed
if (!require(tidyr)) {
  install.packages("tidyr")
}
if (!require(dplyr)) {
  install.packages("dplyr")
}

library(tidyr)
library(dplyr)


# Load the CSV file, skipping the first row and treating the first valid row as the header
file_path <- "/Users/chunsheng/Desktop/Ratio Of Female To Male Youth Unemployment Rate.csv"
data <- read_csv(file_path, skip = 2)
print(colnames(data))
# Check the structure of the loaded data
str(data)

# Show the first few rows
head(data)
# Reshaping the data: converting year columns into rows
# Reshaping the data: converting year columns into rows
tidy_data <- data %>%
  pivot_longer(
    cols = starts_with("19") | starts_with("20"),  # Select year columns
    names_to = "Year",                             # Create 'Year' column
    values_to = "Value"                            # Create 'Value' column
  ) %>%
  rename(
    `Country Name` = `Country Name`,
    `Country Code` = `Country Code`,
    `Indicator Name` = `Indicator Name`,
    `Indicator Code` = `Indicator Code`
  )

# Display the first few rows of the reshaped data
head(tidy_data)


# Select only the "Country Name" and "Country Code" columns
countryname_tibble <- tidy_data %>%
  select(`Country Name`, `Country Code`) %>%
  distinct()  # Fixing the parentheses here

# Convert to a tibble
country_name <- as_tibble(countryname_tibble)

# View the resulting tibble
print(country_name)  # Print the tibble to see the output


# Select only the "Country code", "year" and "value" columns
indicatordata_tibble <- tidy_data %>% select("Country Code", "Year", "Value") %>%
distinct()
# Convert to a tibble
indicator_data <- as_tibble(indicatordata_tibble)
# View the resulting tibble
indicator_data



```

# Problem 5
```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)



# Filter the data for the last 20 years
last_20_years <- 2020 - 20
filtered_data <- tidy_data %>%
  filter(Year >= last_20_years)

# Select countries from each continent
north_america_countries <- c("USA", "CAN", "MEX", "CUB", "JAM")
asia_countries <- c("CHN", "IND", "JPN", "KOR", "IDN")
middle_east_countries <- c("ARE", "SAU", "ISR", "IRN", "EGY")

# Create a vector to categorize the countries
filtered_data <- filtered_data %>%
  mutate(Continent = case_when(
    `Country Code` %in% north_america_countries ~ "North America",
    `Country Code` %in% asia_countries ~ "Asia",
    `Country Code` %in% middle_east_countries ~ "Middle East",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(Continent) & !is.na(Value))

# Plotting the data
ggplot(filtered_data, aes(x = Year, y = Value, color = `Country Name`, group = `Country Name`)) +
  geom_line() +
  labs(title = "Ratio of Female to Male Youth Unemployment Rate",
       x = "Year",
       y = "Ratio (Female to Male)",
       color = "Country") +
  facet_wrap(~ Continent) +
  theme_minimal()
```

##Asia: The ratio of female to male youth unemployment is significantly high in several countries, particularly China and Indonesia. This suggests that young women in these countries face greater challenges in accessing employment compared to their male counterparts.
##Middle East: Countries like Egypt and Iran show substantial fluctuations in their ratios, indicating instability in female youth employment rates over the past two decades. This might reflect socio-political dynamics affecting women’s participation in the labor force.
##North America: The ratios for the United States and Canada are notably low and stable, suggesting a more balanced employment scenario for young men and women. This may reflect effective policies promoting gender equality in the job market.
#Comparative Insights:

#Higher Ratios: Countries with ratios above 100 indicate that young women are significantly disadvantaged in the job market compared to men. For example, the ratios for Egypt and China exceed 200 at times, signaling a critical need for interventions aimed at enhancing female employment.
#Lower Ratios: Countries like Canada and the United States maintain ratios under 100, suggesting relatively equitable youth unemployment conditions. This could be attributed to better access to education and employment opportunities for young women.




