---
title: "DA5020.A6.Chunsheng.Fang"
output: 
  pdf_document: default
  html_document: default
date: "2024-10-20"
---


#Connect to database
```{r}
# Load necessary libraries
library(DBI)
library(RSQLite)

# Set working directory (where your SQLite database and CSV file are located)
setwd("/Users/chunsheng/sqlite")  # Update with your actual path

sqlite <- dbDriver("SQLite")
conn <- dbConnect(sqlite,"imdb.db")




```

#Check the tables in the database
```{r}
as.data.frame(dbListTables(conn))
```
#Check the tables in the database
```{r}
dbGetQuery(conn, "SELECT AVG(Runtime) AS avg_runtime
FROM movie_info
WHERE Genre LIKE '%thriller%'")
```
# Problem 1


OPEN DATABASE
sqlite3 imdb.db

.open imdb.db

#1.Create directors table

CREATE TABLE director_info (
Director_ID INTEGER PRIMARY KEY,
Director_Name TEXT );


#2.Import directors csv to the

.mode csv

.import directors.csv director_info

SELECT * FROM director_info LIMIT 5;

There was a datatype mismatch error, hence I changed director_ID datatype from INT to Numeric

ALTER TABLE director_info MODIFY COLUMN Director_ID Numeric;

# Problem 2
```{r}
library(ggplot2)
#1 Count the number of rows in the movie_info and director_info tables
dbGetQuery(conn, "SELECT COUNT(*) as row_count FROM movie_info")

dbGetQuery(conn, "SELECT COUNT(*) as row_count FROM director_info")

# 2 How many movies were released between 2010 and 2020 (inclusive)

dbGetQuery(conn, "SELECT COUNT(Release_Year) AS num_movies
FROM movie_info
WHERE Release_Year BETWEEN '2010' AND '2020'")

num_movies <- dbGetQuery(conn, "SELECT Release_Year, COUNT(Release_Year) AS num_movies
FROM movie_info
WHERE Release_Year BETWEEN '2010' AND '2020'
GROUP BY Release_Year")

ggplot(num_movies, aes(x = Release_Year, y = num_movies, fill = Release_Year)) +
geom_bar(stat = "identity", width = 0.5) +
theme_minimal() +
labs(x = "Year", y = "Number of movies", fill = "Year") +
scale_fill_discrete(name = "Year", labels = num_movies$Release_Year) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


#3 What is the minimum, average and maximum ratings for “Action” movies.

dbGetQuery(conn, "SELECT MIN(IMDB_Rating) AS min_rating, AVG(IMDB_Rating) AS avg_rating, MAX(IMDB_Rating) AS max_rating
FROM movie_info
WHERE Genre LIKE '%Action%'")


#4 What are the 25 highest-grossing movies within the dataset? 
dbGetQuery(conn, "SELECT Series_Title, Genre, CAST(Gross AS REAL) AS Gross
FROM movie_info
WHERE Gross IS NOT NULL
ORDER BY Gross DESC
LIMIT 25")



#5 Which directors have the highest-grossing movies.
Totalgross <- dbGetQuery(conn, "SELECT d.Director_Name, SUM(m.Gross) AS Total_Gross
FROM movie_info m
JOIN director_info d ON m.Director_ID = d.Director_ID
GROUP BY d.Director_Name
ORDER BY Total_Gross DESC
LIMIT 10")
Totalgross


ggplot(Totalgross, aes(x = Director_Name, y = Total_Gross)) +
geom_bar(stat = "identity") +
theme_minimal()+
xlab("Director Name") +
ylab("Total Gross") +
ggtitle("Directors with the highest-grossing movies")+
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


#6 Create a function called verifyDirector() t
verifyDirector <- function(director_name) {
# Query the database for the director name
query <- dbGetQuery(conn, paste("SELECT Director_Name
FROM director_info
WHERE Director_Name = '", director_name, "'", sep=""))
# Check if the director was found
if (nrow(query) == 0) {
message(paste("Director", director_name, "not found"))
} else {
message(paste("Director", director_name, "found"))
} }
# Call the verifyDirector() function with a director name argument
verifyDirector("Anurag Basu")

verifyDirector("Apeksha udayakumar")

```




